---
title: "Facets"
output: 
  html_document:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

## Agenda

* Basic faceting in `ggplot2`
  
* When to free scales

* Getting data in the right form

* Plotting one variable per panel

* Design decisions with multiple variables


## Without faceting

```{r}
library(tidyverse)
g <- ggplot(iris, aes(Sepal.Length, Sepal.Width)) +
  geom_point()
g
```

## Facet on one variable with `facet_wrap()`

facet "on" Species

```{r}
g +
  facet_wrap(~Species)
```

## Faceting in ggplot2

Each panel represents one categorical group / levels of a factor (type can be factor or character or integer)

```{r}
glimpse(iris)
```


## Change the layout with `nrow, ncol`

```{r}
g +
  facet_wrap(~Species, ncol = 1)
```

## Add regression lines


```{r}
g +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~Species)
```

## Facet on two (categorical) variables with `facet_grid()`

```{r}
ggplot(mtcars, aes(hp,mpg)) +
  geom_point() +
  facet_grid(cyl~gear)
```

## Label variables (in addition to factor levels)

```{r}
ggplot(mtcars, aes(hp, mpg)) +
  geom_point() +
  facet_grid(cyl~gear, labeller = label_both)
```


## Cleveland dot plots

Note that y-axis is discrete

```{r}
g <- mtcars |>
  rownames_to_column("car") |>
  ggplot(aes(mpg, reorder(car, mpg))) +
  geom_point(color = "blue") +
  ylab("") +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
g
```

## Facet by gear

```{r}
g +
  facet_grid(gear ~ ., labeller = label_both)
```

## "Free" the y scale with `scales = "free_y"`

```{r}
g +
  facet_grid(gear ~ ., labeller = label_both,
             scales = "free_y")
```

## Change panel heights with `space = "free_y"`

```{r}
g +
  facet_grid(gear ~ ., labeller = label_both,
             scales = "free_y", space = "free_y")
```

## In general, do not "free" numerical scales

```{r}
ggplot(iris, aes(Sepal.Length, Sepal.Width)) +
  geom_point() +
  facet_wrap(~Species, scales = "free")
```

incorrect

## In general, do not "free" numerical scales

```{r}
ggplot(iris, aes(Sepal.Length, Sepal.Width)) +
  geom_point() +
  facet_wrap(~Species)
```

correct

## EXERCISE

Try faceting with `ggplot2movies` dataset