---
title: "Scales"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
```

## Problematic y-axis

```{r}
library(tidyverse)
ggplot(mtcars, aes(x = hp)) +
  geom_histogram(breaks = seq(0, 400, 50),
                 color = "blue", fill = "cornflowerblue")
```

## Change scale breaks

```{r}
ggplot(mtcars, aes(x = hp)) +
  geom_histogram(breaks = seq(0, 400, 50),
                 color = "blue", fill = "cornflowerblue") +
  scale_y_continuous(breaks = seq(0, 10, 2))
```

## Change axis limits

```{r}
ggplot(mtcars, aes(x = hp)) +
  geom_histogram(breaks = seq(0, 400, 50),
                 color = "blue", fill = "cornflowerblue") +
  scale_y_continuous(limits = c(0, 15))
```

## Problematic tick mark labels

```{r}
ggplot(mtcars, aes(x = factor(am), y = mpg)) + 
  geom_boxplot()
```


## Change tick mark labels

```{r}
ggplot(mtcars, aes(x = factor(am), y = mpg)) + 
  geom_boxplot() + 
  scale_x_discrete(name = "transmission type",
                   labels = c("0" = "automatic", "1" = "manual"))
```

## Discrete vs. continuous scales (pitfall alert)

```{r}
#| warning: true
#| error: true
ggplot(mtcars, aes(x = factor(am), y = mpg)) + 
  geom_boxplot() + 
  scale_x_continuous(name = "transmission type",
                   labels = c("0" = "automatic", "1" = "manual"))
```


## Discrete vs. continuous scales (pitfall alert)

```{r}
#| warning: true
#| error: true
ggplot(mtcars, aes(x = factor(am), y = mpg)) + 
  geom_boxplot() + 
  scale_y_discrete(limits = c(0, 40))

```

## Discrete color scales (manual)

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(am))) +
  geom_point() +
  scale_color_manual(values = c("#FF8146", "#009BFF"))
```

## Discrete color scales (prebuilt)

```{r}
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(am))) +
  geom_point() +
  scale_color_brewer(palette = "Set1")
```

## Discrete color scales (prebuilt)

```{r}
library(ggthemes)
ggplot(mtcars, aes(x = wt, y = mpg, color = factor(am))) +
  geom_point() +
  scale_color_colorblind()
```

## Mixing up color and fill

```{r}
ggplot(mtcars, aes(x = wt, y = mpg,
                   color = factor(am))) +
  geom_point() +
  scale_fill_manual(values = c("orange", "black"))
```

## Coordinate systems

```{r}
ggplot(mtcars, aes(x = mpg, y = factor(gear))) +
  geom_boxplot() +
  coord_flip()
```

## ... but scales stay with original mappings

```{r}
ggplot(mtcars, aes(x = mpg, y = factor(gear))) + geom_boxplot() + 
  scale_x_continuous(name = "Miles per gallon") + 
  coord_flip()
```

## Better

```{r}
ggplot(mtcars, aes(x = factor(gear), y = mpg)) +
  geom_boxplot() +
  scale_y_continuous(name = "Miles per gallon")
```

